% !TeX root = ../../infdesc.tex
\section{Set operations}
\secbegin{secSetOperations}

\begin{exercise}
Prove that $[0,\infty) \cap \mathbb{Z} = \mathbb{N}$.
\end{exercise}

\begin{exercise}
Write down the elements of the set
\[ \{ 0, 1, 4, 7 \} \cap \{ 1, 2, 3, 4, 5 \} \]
\end{exercise}

\begin{exercise}
Express $[-2,5) \cap [4,7)$ as a single interval.
\end{exercise}


\begin{exercise}
Let $X$ be a set. Prove that $X \cap \varnothing = \varnothing$.
\end{exercise}



\begin{exercise}
Let $a,b,c,d \in \mathbb{R}$ with $a<b$ and $c<d$. Prove that the open intervals $(a,b)$ and $(c,d)$ are disjoint if and only if $b<c$ or $d<a$.
\end{exercise}


\begin{exercise}
Write down the elements of the set
\[ \{ 0, 1, 4, 7 \} \cup \{ 1, 2, 3, 4, 5 \} \]
\end{exercise}

\begin{exercise}
Express $[-2,5) \cup [4,7)$ as a single interval.
\end{exercise}


\begin{exercise}
Let $X$ and $Y$ be sets. Prove that $X \subseteq Y$ if and only if $X \cup Y = Y$.
\end{exercise}



\begin{exercise}
\label{exUnionDistributesOverIntersection}
Let $X,Y,Z$ be sets. Prove that $X \cup (Y \cap Z) = (X \cup Y) \cap (X \cup Z)$.
\end{exercise}


\begin{exercise}
Express $\displaystyle\bigcup_{n \ge 1} [0,1-\frac{1}{n}]$ as an interval.
\end{exercise}

\begin{exercise}
Prove that $\displaystyle\bigcap_{n \in \mathbb{N}} [n] = \varnothing$ and $\displaystyle\bigcup_{n \in \mathbb{N}} [n] = \{ k \in \mathbb{N} \mid k \ge 1 \}$.
\end{exercise}


\begin{exercise}
\label{exIndexedIntersectionUnionGeneralisePairwise}
Let $X_1$ and $X_2$ be sets. Prove that
\[ X_1 \cap X_2 = \bigcap_{k \in [2]} X_k \quad \text{and} \quad X_1 \cup X_2 = \bigcup_{k \in [2]} X_k \]
\end{exercise}

\begin{exercise}
\label{exSubsetsFiniteIntersectionInhabitedInfiniteIntersectionEmpty}
Find a family of sets $\{ X_n \mid n \in \mathbb{N} \}$ such that:
\begin{enumerate}[(i)]
\item $\displaystyle\bigcup_{n \in \mathbb{N}} X_n = \mathbb{N}$;
\item $\displaystyle\bigcap_{n \in \mathbb{N}} X_n = \varnothing$; and
\item $X_i \cap X_j \ne \varnothing$ for all $i,j \in \mathbb{N}$.
\end{enumerate}
\begin{backhint}
\hintref{exSubsetsFiniteIntersectionInhabitedInfiniteIntersectionEmpty}
You need to find a family of subsets of $\mathbb{N}$ such that (i) any two of the subsets have infinitely many elements in common, but (ii) given any natural number, you can find one of the subsets that it is \textit{not} an element of.
\end{backhint}
\end{exercise}


\begin{exercise}
Write down the elements of the set
\[ \{ 0, 1, 4, 7 \} \setminus \{ 1, 2, 3, 4, 5 \} \]
\end{exercise}

\begin{exercise}
Express $[-2,5) \setminus [4,7)$ and $[4,7) \setminus [-2,5)$ as intervals.
\end{exercise}

\begin{exercise}
\label{exSetMinusSetMinus}
Let $X$ and $Y$ be sets. Prove that $Y \setminus (Y \setminus X)= X \cap Y$, and deduce that $X \subseteq Y$ if and only if $Y \setminus (Y \setminus X) = X$.
\end{exercise}


\begin{exercise}
Complete the proof of de Morgan's laws for sets.
\end{exercise}


\begin{exercise}
Write down the elements of the set $\{ 1, 2 \} \times \{ 3, 4, 5 \}$.
\end{exercise}

\begin{exercise}
Let $X$ be a set. Prove that $X \times \varnothing = \varnothing$.
\end{exercise}

\begin{exercise}
\label{exCartesianProductNotAssociative}
Let $X$, $Y$ and $Z$ be sets. Under what conditions is it true that $X \times Y = Y \times X$? Under what conditions is it true that $(X \times Y) \times Z = X \times (Y \times Z)$?
\end{exercise}
