% !TeX root = ../../infdesc.tex
\section{Relations}
\secbegin{secRelations}


\begin{exercise}
Give three more examples of relations, not all of which are homogeneous.
\end{exercise}



\begin{exercise}
Let $R$ and $S$ be relations on $\mathbb{R}$ defined for $a,b \in \mathbb{R}$ by letting
\[ a \mathrel{R} b ~ \Leftrightarrow ~ b-a \in \mathbb{Q} \quad \text{and} \quad a \mathrel{S} b ~ \Leftrightarrow ~ \exists n \in \mathbb{Z},~ (n \ne 0) \wedge n(b-a) \in \mathbb{Z} \]
Prove that $R=S$.
\end{exercise}


\begin{exercise}
Let $R$ be the relation on $\mathbb{Z}$ defined for $x,y \in \mathbb{Z}$ by letting $x \mathrel{R} y$ if and only if $x^2=y^2$. Describe its graph $\mathrm{Gr}(R) \subseteq \mathbb{Z} \times \mathbb{Z}$.
\end{exercise}

\begin{exercise}
Let $f : X \to Y$ be a function, and define the relation $R_f$ from $X$ to $Y$ as in \Cref{exRelationExamples}. Prove that $\mathrm{Gr}(R_f) = \mathrm{Gr}(f)$---that is, the graph of the \textit{relation} $R_f$ is equal to the graph of the \textit{function} $f$.
\end{exercise}


\begin{exercise}
\label{exGraphOfEmptyAndDiscreteRelations}
Let $X$ and $Y$ be sets. Describe the graphs $\mathrm{Gr}(D_{X,Y})$ and $\mathrm{Gr}(\varnothing_{X,Y})$.
\end{exercise}


\begin{exercise}
Let $X$ be a set. Prove that $\Delta_X = \mathrm{Gr}(=)$.
\end{exercise}



\begin{exercise}
\label{exSubsetIsReflexive}
Let $X$ be a set. Prove that $\subseteq$ is a reflexive relation on $\mathcal{P}(X)$, but $\subsetneqq$ is not.
\end{exercise}

\begin{exercise}
\label{exDivisibilityIsReflexive}
Prove that the relation `$x$ divides $y$' on $\mathbb{Z}$ is reflexive.
\end{exercise}


\begin{exercise}
\label{exReflexivitySensitiveToDomainOfRelation}
Let $G = \{ (1,1), (2,2), (3,3) \}$. Let $R$ be the relation on $[3]$ whose graph is $G$, and let $S$ be the relation on $[4]$ whose graph is $G$. Prove that $R$ is reflexive, but $S$ is not.
\end{exercise}



\begin{exercise}
Find all subsets $U \subseteq \mathbb{Z}$ such that the relation `$x$ divides $y$' on $U$ is symmetric.
\end{exercise}


\begin{exercise}
\label{exSymmetryNotSensitiveToDomainOfRelation}
Let $R$ and $S$ be relations such that $\mathrm{Gr}(R) = \mathrm{Gr}(S)$. Note that the domain of $R$ might be different from the domain of $S$. Prove that $R$ is symmetric if and only if $S$ is symmetric.
\end{exercise}


\begin{exercise}
\label{exDivisibilityIsOrIsNotAntisymmetric}
Prove that the relation `$x$ divides $y$' on $\mathbb{N}$ is antisymmetric, but not on $\mathbb{Z}$.
\end{exercise}

\begin{exercise}
Let $X$ be a set. Prove that $\subseteq$ is an antisymmetric relation on $\mathcal{P}(X)$.
\end{exercise}

\begin{exercise}
Let $X$ be a set and let $R$ be a relation on $X$. Prove that $R$ is both symmetric and antisymmetric if and only if $\mathrm{Gr}(R) \subseteq \Delta_X$, where $\Delta_X$ is the diagonal subset of $X \times X$ (see \Cref{defDiagonalSubset}). Deduce that the only reflexive, symmetric and antisymmetric relation on a set $X$ is the equality relation on $X$.
\end{exercise}


\begin{exercise}
Let $X$ be a set. Prove that $\subseteq$ is a transitive relation on $\mathcal{P}(X)$.
\end{exercise}

\begin{exercise}
\label{exDivisibilityIsTransitive}
Prove that the relation `$x$ divides $y$' on $\mathbb{Z}$ is transitive.
\end{exercise}


\begin{exercise}
\label{exTransitivityNotSensitiveToDomainOfRelation}
Let $R$ and $S$ be relations such that $\mathrm{Gr}(R) = \mathrm{Gr}(S)$. Note that the domain of $R$ might be different from the domain of $S$. Prove that $R$ is transitive if and only if $S$ is transitive.
\end{exercise}
